Presentation Title

### 2017-06-07 thesis/chapters/ch2-model.tex
Write validation

% The TX protocol~\footnote{It is also known as True Halves.
% See \url{https://github.com/Tribler/tribler/pull/2135} for more information.}
% consist of two messages, \texttt{tx\_req} and \texttt{tx\_resp}.
% When a node  wishes to create a transaction,
% it generates a TX block with the corresponding hash pointer, sequence number, counterparty and the transaction message itself.
% It then sends a \texttt{tx\_req} message containing the TX block to the counterparty.
% Upon receiving a TX block at the counterparty, it creates its own version of the TX block (the pair of the sender's TX block),
% and sends it back to the sender in a \texttt{tx\_resp} message.
% At the end of the protocol, both parties should have both versions of the TX block.

% \subsection{Validation Protocol}
% The consensus and transaction protocol by themselves do not provide a mechanism to detect forks or other forms of tamperaing.
% This is the goal of the validation protocol.

% It also has two message---\texttt{vd\_req} and \texttt{vd\_resp}.
% When a node wishes to validate a transaction,
% it sends a \texttt{vd\_req} with the sequence number $s$ to the party that created the TX block of the transaction.
% Upon receiving a \text{vd\_req}, the node finds two \emph{agreed} CP blocks that surrounds the TX block and computes the \emph{fragment}.
% An agreed CP block means that it has reached consensus.
% A fragment of two CP blocks is a section of the chain where the beginning and the end are defined by those two CP blocks.
% With the fragment, the node replies with a \texttt{vd\_resp} message.
% Upon receiving a \texttt{vd\_resp}, the node checks that the fragment is created correctly.
% Loosely, that means the agreed CP blocks have actually reached consensus,
% the fragment has the correct hash pointers,
% and TX block containing $s$ exists and is valid.

@@

For notational clarify, we use the following convention (adapted from~\cite{miller2016honey}) throughout this work.
\begin{itemize}
\item Lower case (e.g. $x$) denotes a scalar object.
\item Upper case (e.g. $X$) denotes a set or a constant.
\item Sans serif (e.g. $\textsf{fn}(\cdot)$) denotes a function.
\item Typewrite (e.g. $\texttt{ack}$) denotes message type.
\end{itemize}

@@

\begin{algorithm}
\caption{Function $\textsf{derive\_facilitator}(C, n)$ takes a list of CP blocks $C$ and an integer $n$,
sort evey element in $C$ by its luck value (the $\lambda$-expression), and outputs the smallest $n$ elements.}
\label{alg:facilitator}
\begin{algorithmic}
\State $\textbf{take} (n, \textbf{sort} (\textbf{map}(\lambda x.\textsf{H}(x || pk \text{ of x}), C)))$
\end{algorithmic}
\end{algorithm}

@@

@@

The TX protocol, shown in \Cref{alg:tx-proto}, is run by all nodes.
It is also known as True Halves TODO.
Node that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
$txid$ should be a uniformly distributed random value, i.e. $txid \in_R \{0, 1\}^{256}$.
Then the initiator sends $\langle \texttt{tx\_req}, t_{u, h}\rangle$ to $v$.

@@

\begin{algorithm}
    \caption{The TX protocol which runs in the context of node $u$.}
    \label{alg:tx-proto}

    \begin{algorithmic}
        \Upon $\langle \texttt{tx\_req}, t_{v, j} \rangle$
        \State $txid, pk_v, m \gets t_{v, j}$ \Comment Unpack the TX
        \State $\textsf{new\_tx}(pk_u, m, txid)$
        \State store $t_{v, j}$ as the other half of $t_{u, h}$
        \State send $\langle \texttt{tx\_reqp}, t_{u, h} \rangle$ to $v$

        \Upon $\langle \texttt{tx\_resp}, t_{v, j} \rangle$
        \State $txid, pk_v, m \gets t_{v, j}$ \Comment Unpack the TX
        \State store $t_{v, j}$ as the other half of the TX with identifier $txid$
    \end{algorithmic}
\end{algorithm}

A key feature of the TX protocol is that it is non-blocking.
TODO

@@

>>>>

@@

For notational clarify, we use the following convention (adapted from~\cite{miller2016honey}) throughout this work.
\begin{itemize}
\item Lower case (e.g. $x$) denotes a scalar object or a tuple.
\item Upper case (e.g. $X$) denotes a set or a constant.
\item Sans serif (e.g. $\textsf{fn}(\cdot)$) denotes a function.
\item Typewrite (e.g. $\texttt{ack}$) denotes message type.
\end{itemize}

@@

\begin{algorithm}
\caption{Function $\textsf{derive\_facilitator}(C, n)$ takes a list of CP blocks $C$ and an integer $n$,
sort evey element in $C$ by its luck value (the $\lambda$-expression), and outputs the smallest $n$ elements.}
\label{alg:facilitator}
\begin{algorithmic}
\State $\textbf{take} (n, \textbf{sort} (\textbf{map}(\lambda x.\textsf{H}(x || pk \text{ of } x), C)))$
\end{algorithmic}
\end{algorithm}

@@

TODO assume the wait time is much longer than network delay.

@@

The TX protocol, shown in \Cref{alg:tx-proto}, is run by all nodes.
It is also known as True Halves, first described by Veldhuisen~\cite[Chapter~3.2]{truehalves}.
Node that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
$txid$ should be a uniformly distributed random value, i.e. $txid \in_R \{0, 1\}^{256}$.
Then the initiator sends $\langle \texttt{tx\_req}, t_{u, h}\rangle$ to $v$.

@@

\begin{algorithm}
    \caption{The TX protocol which runs in the context of node $u$. TODO}
    \label{alg:tx-proto}

    \begin{algorithmic}
        \Upon $\langle \texttt{tx\_req}, t_{v, j} \rangle$ from $v$
        \State $txid, pk_v, m \gets t_{v, j}$ \Comment unpack the TX
        \State $\textsf{new\_tx}(pk_u, m, txid)$
        \State store $t_{v, j}$ as the pair of $t_{u, h}$
        \State send $\langle \texttt{tx\_reqp}, t_{u, h} \rangle$ to $v$

        \Upon $\langle \texttt{tx\_resp}, t_{v, j} \rangle$ from $v$
        \State $txid, pk_v, m \gets t_{v, j}$ \Comment unpack the TX
        \State store $t_{v, j}$ as the pair of the TX with identifier $txid$
    \end{algorithmic}
\end{algorithm}

A key feature of the TX protocol is that it is non-blocking.
At no time in \Cref{alg:new-tx} or \Cref{alg:tx-proto} do we need to hold the chain state and wait for some message to be delivered before committing a new block to the chain.
This allows for high concurrency where we can call $\textsf{new\_tx}(\cdot)$ multiple times without waiting for the corresponding \texttt{tx\_resp} messages.

@@

Up to this point, we do not provide a mechanism to detect forks or other forms of tampering or forging, and we cannot detect malicious parties.
The validation protocol aims to solve this issue.
The protocol is also a request-response protocol, just like the transaction protocol.
But before explaining the protocol itself, we first define what it means to be valid.

\subsection{Validity Definition}
A \emph{valid} TX must be created by following the TX protocol.
Concretely, a TX $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), i, txid_u, pk_v, m_u, sig_u \rangle$ should satisfy the following conditions.
\begin{enumerate}
    \item \emph{Existance of pair}:
        There exists $t_{v, j} = \langle \textsf{H}(b_{v, j - 1}), j, txid, pk_u, m, sig_v \rangle$,
        i.e. the pair, where $txid_u = txid_v$, $m_u = m_v$ and the signatures are valid.
    \item \emph{Existance of fragment}:
        There exists $\textsf{agreed\_fragment}(t_{u, i})$---call it $F_{u, i}$ and
        $\textsf{agreed\_fragment}(t_{v, j})$---call it $F_{v, j}$,
        where the hash pointers are valid and the enclosing CP blocks have valid signatures.
    \item \emph{Only one pair}:
        There is only one transaction with $txid$ in $F_{v, j}$.
    \item \emph{Small fragment}:
        The $|F_{u, i}| \le L$ and $|F_{v, j}| \le L$.
    \item \emph{Same consensus round}:
        Suppose the last block in $F_{u,i}$ and the last block in $F_{v, j}$ are respectively $c_{u, p}$ and $c_{v, q}$. 
        Then these they must be created from the same consensus round.
\end{enumerate}
We make three remarks regarding the definition.
Firstly, the existance of fragment is the most important property.
It ensures that the pair of TX blocks exist between a pair of CP blocks that are in consensus.
This implicitly proves that the TX blocks are in consensus (we show a proof in \Cref{sec:analysis} TODO).
Secondly, the same consensus round property is to encourage node to remain active in the network.
Recall that the facilitators waits for duration $\Delta$, which is much longer than the network delay.
Thus the \texttt{cp\_msg} should be delivered to the facilitators before the consensus starts.
Lastly, TX blocks that do not satisfy the requirement above are not invalid.
In fact, most new TX blocks cannot satisfy the Existance of fragment property because the CP blocks following the new transactions have not reached consensus.
Moreoever, very new TX blocks may not satisfy the Existance of pair property because the counterparty may have not yet sent \texttt{tx\_resp}.
Next we define \emph{invalid} transactions.
\begin{enumerate}
    \item \emph{Bad pair}:
        $txid_u = txid_v \wedge m_u \ne m_v$.
    \item \emph{Invalid counterparty}:
        $txid_u = txid_v$, but $pk_u$ in $t_{v, j}$ does not match $u$'s public key.
    \item \emph{Invalid hash pointer}:
        Suppose $F_{u, i}$ and $F_{v, j}$ are available, but the transactions in the fragments do not form a chain.
    \item \emph{Large fragment}:
        $F_{u, i} > L \vee F_{v, j} > L$.
    \item \emph{Different consensus round}:
        The last block in $F_{u,i}$ and the last block in $F_{v, j}$ are respectively $c_{u, p}$ and $c_{v, q}$. 
        But these CP blocks are not created from the same consensus round.
\end{enumerate}
The final two properties are the negation of the final two properties in the valid definition.
The other three properties are for detecting tampering.
Any transaction that do not fall into the valid or invalid category are known as \emph{unknown} transactions.

\subsection{Validation Protocol}
With the validity definition, we are ready to construct a protocol for determining the validity of transactions.
The protocol is a simple response and request protocol (\Cref{alg:vd-proto}).
If $u$ wishes to validate some TX with ID $txid$ and counterparty $v$, it sends $\langle \texttt{vd\_req}, txid \rangle$ to $v$.
The desired properties of the validation protocol are as follows.
\begin{enumerate}
    \item \emph{Correctness}:
        The validation protocol outputs the correct result
        according to the aforementioned validity definition.
    \item \emph{Agreement}:
        If any correct node decides on the validity of a transaction,
        then all other correct nodes are able to reach the same conclusion.
    \item \emph{Liveness}:
        Any valid transactions can be validated eventually.
    \item \emph{Unforgeability}:
        If some transaction is valid, it cannot be forged into an invalid transaction.
        If some transaction is invalid, it cannot be forged into a valid transaction.
\end{enumerate}

\begin{algorithm}
\caption{Validation protocol}
\label{alg:vd-proto}

\begin{algorithmic}
    \Upon $\langle \texttt{vd\_req}, txid \rangle$ from $v$
        \State $t_{u, i} \gets \text{the transaction identified by } txid$
        \State $F_{u, i} \gets \textsf{agreed\_fragment}(t_{u, i})$
        \State send $\langle \texttt{vd\_resp}, txid, F_{u, i} \rangle$ to $v$

    \Upon $\langle \texttt{vd\_resp}, txid, F_{v, j} \rangle$ from $v$
        \State $t_{u, i} \gets \text{the transaction identified by } txid$
        \State set the validity of $t_{u, i}$ to $\textsf{get\_validity}(t_{u, i}, F_{v, j})$
\end{algorithmic}
\end{algorithm}

The core of the validation protocol lies within the $\textsf{get\_validity}(\cdot)$ function (\Cref{alg:get-validity}).
It essentially implements all the aforementioned validity requirements.
If they are all satisfied, the function returns \emph{valid}.
Otherwise, depending on which stage the function failed, it will return either \emph{unknown} or \emph{invalid}.
We prove in the following chapter that the validation protocol satisfies the most of the desired properties.

\begin{algorithm}
\caption{Function $\textsf{get\_validity}(t_{u, i}, F_{v, j})$ runs in the private context of $u$.
$t_{u, i}$ is the transaction that $u$ wishes to verify, and $F_{v, j}$ is the corresponding fragment received from $v$.}
\label{alg:get-validity}

\begin{algorithmic}

    \State $F_{u, i} \gets \textsf{agreed\_fragment}(t_{u, i})$
    \If{$F_{u, i} = \bot$}
        \State \Return \emph{unknown}
    \EndIf
    \State

    \State $c_{v, a} \gets \textsf{first}(F_{v, j})$
    \State $c_{v, b} \gets \textsf{last}(F_{v, j})$
    \If{$c_{v, a}$ or $c_{v, b}$ are not in consensus}
        \State \Return \emph{unknown}
    \EndIf
    \State

    \State $txid, pk_v, m \gets t_{u, i}$
    \If{number of blocks of $txid$ in $F_{v, j} = 0$}
        \State \Return \emph{unknown}
    \EndIf
    \State

    \If{number of blocks of $txid$ in $F_{v, j} > 1$}
        \State \Return \emph{invalid}
    \EndIf
    \State

    % txid are equal now, but bad hash pointer
    \If{sequence number in $F_{v, j}$ is correct (sequential)}
        \If{hash pointers in $F_{v, j}$ is wrong}
            \State \Return \emph{invalid}
        \EndIf
    \EndIf
    \State

    \State $txid', pk'_u, m' \gets t_{v, j}$
    \If{$m \ne m' \vee pk_u \ne pk_u \vee |F_{v, j}| > L$}
        \State \Return \emph{invalid}
    \EndIf
    \State

    \State $c_{u, b} \gets \textsf{last}(F_{u, i})$
    \If{$c_{u, b}$ is not created using the same $\C_r$ as $c_{v, b}$}
        \State \Return \emph{invalid}
    \EndIf
    \State

    \State \Return \emph{valid}
\end{algorithmic}
\end{algorithm}

We make two remarks.
First, just like the TX protocol, we do not block at any part of the protocol.
Second, suppose some $F_{v, j}$ validates $t_{u, i}$, then that does not imply that $t_{u, i}$ only has one pair $t_{v, j}$.
Our validity requirement only requires that there is only one $t_{v, j}$ in the correct consensus round.
The counterparty may create any number of fake pairs in a later consensus rounds.
But these fake pairs only pollutes the chain of $v$ and can never be validated because the round is incorrect.
